\documentclass{article}
\usepackage[utf8]{inputenc}

\title{ECON 5253 PS1 Overleaf Test}
\author{William Lorton}
\date{February 2021}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Profile}

My primary reason for taking this course is to further develop my programming skills in R and also pick up on Python, SQL, and anything else that may be of use. I plan to supplement any coding work we do with practice sessions via something such as DataCamp. I'm also interested in learning more about Machine Learning as  I have had very little experience with it. I will probably read through the \emph{Master Algorithm} book that you mentioned before (the first chapter was entertaining at least); I think DataCamp may have some courses on ML as well.

\medskip

I'm not currently set on a research idea. I will likely peruse Kaggle or the GSS website for a data set that seems viable. I think something involving sentiment analysis could be interesting (e.g. of Reddit posts in light of the recent meme stock havoc), though I'm not sure how I would use it at this point.

\medskip

As for my plans after graduation, I'm looking to apply to jobs that are focused on research of some sort (ideally such that the skills taught in this class can be leveraged in one way or another). As an example, I'd be interested in data analyst/scientist positions as well as financial analyst positions (particularly those involved with investing). 

\medskip

Thanks for pointing out this website by the way. It seems like an efficient way to learn LaTeX and produce elegant documents.

\section{Equation}

Test equation:

\begin{displaymath}
a^2 + b^2 = c^2 
\end{displaymath}


\end{document}
