\documentclass{homeworg}

\title{ECON 5253: PS2}
\author{William Lorton}

\begin{document}

\maketitle

\section*{Data Scientist Tools List}

\begin{enumerate}
   \item Measurement
   \begin{itemize}
     \item Must know how to appropriately measure the components involved in your task.
     \item Often involves giving a quantitative interpretation to a problem that began with only a qualitative dimension.
     \item Leads to concrete policy creation.
   \end{itemize}
   \item Statistical programming languages
   \begin{itemize}
    \item Main DS languages are R, Python, and Julia.
    \item Stata, SAS, SPSS, Matlab, and JavaScript are also used.
    \item R and Python are the most popular currently (Julia is the relatively new up-and-comer).
    \item Note the difference between scripted languages (e.g. those listed above) and faster compiled languages like C++, C, and FORTRAN.
   \end{itemize}
   \item Data visualization
   \begin{itemize}
    \item Communicate insights to those with and without intimate knowledge of the data.
    \item Identify outliers, check for correctness, see data from new perspectives, etc.
    \item R: ggplot2 package; Python: matplotlib package; Julia: Plots.jl package.
    \item Lots of companies use interactive software like Tableau as well.
   \end{itemize}
   \item Big Data management software
   \begin{itemize}
       \item For when data is too large for your hard drive(s) and/or overloads your computer's memory when using R/Python/Julia/Matlab/etc.
       \item Use Resilient Distributed Datasets (RDDs) via Hadoop or Spark; divides data into smaller bits and executes actions on it simultaneously via a cluster of computers.
       \item Can do variety of typical commands: subsetting, merging, summary statistics generation, etc.
       \item Also helpful to know SQL (a database management language) when dealing with very large datasets. 
   \end{itemize}
   \item Data collection tools
   \begin{itemize}
       \item For example, know how to web scrape (lift data off of public web pages very quickly).
       \item  Use an Application Program Interface (API) if available and/or parse HTML file text.
   \end{itemize}
   \item Modeling
   \begin{itemize}
       \item Combine knowledge of statistics and statistical programming languages to test theories, predict outcomes, and establish causal relationships.
       \item Of course, must have already collected, cleaned, and (usually) visualized the data.
   \end{itemize}
\end{enumerate}


\end{document}
